Dismal Code: Studying the Evolution of Security Bugs

### 2013-02-07 docs/msr.tex
more on the methodology and the initial results sections

First, we scanned the Maven repository for appropriate {\sc jar}s.
We discuss the {\sc jar} selection process in the next 
section. With this list we created a series of processing tasks
and added them to the task queue. Then we
executed twenty five (Unix-based) workers that checked out tasks from the queue,
processed the data and stored the results to the data repository.

A typical processing cycle of a worker included the following steps: after
the worker spawns, it requests a task from the queue. This task contains
the {\sc jar} name, which is typically a project version that is downloaded locally.
First, specific {\sc jar} metadata are calculated and stored. Such metadata include the
last modification date of the {\sc jar}, its size and dependencies and a number that
represents the chronological order of the release. This order is derived from an {\sc xml} file that
accompanies every project in the Maven repository called {\it
maven-metadata.xml}. Then FindBugs is executed and its' results are also stored
in the data repository. When the task is completed the queue is notified and
the next task is requested.

@@

Initially, we obtained a snapshot (January 2012) of the maven repository and
handled it locally to retreive a list of all the names of the project versions
that existed in it. Our first problem was that the Maven repository
hosts projects from languages other than Java, typically languages that produce
executable code for the {\sc jvm} (Java Virtual Machine), such as Scala, Groovy,
Clojure etc. These projects were filtered out by a series of checks in
the repository data and metadata.

In addition, we implemented a series of checks in the worker scripts that
checked if the {\sc jar}s are valid in terms of implementation. For instance,
for every {\sc jar} the worker checked if there were any {\it .class} files
before invoking FindBugs. After the project filtering we narrowed down
our data set to 17,055 projects with 114,399 versions.
Table~\ref{tbl:repository} summarises the data set information and
provides the basic descriptive statistic measurements. The distribution of version
count among the selected projects is presented in Figure \ref{fig:version-count}.

@@

The repository has many project with a
few number of versions. On the contrary, there are numerous projects with a
number of versions that is less than ten.
The statistical mesurements presented in Table~\ref{tbl:repository}
indicate that we have 17,505 projects total and the data sets median
is 3, which means that almost 50\% (8,753 projects) of the project
population has 1 to 3 versions. There are a few projects containing ten
versions and only a few with hundreds of versions. The maximum number of
versions for a project is 337. The 3$^{rd}$ quartile measurement
also indicated that 75\% (13,129) of the projects have a maximum of 8 versions.

@@

FindBugs separates software bugs into seven categories (see
Table~\ref{tbl:bug-cat}). Two of them involve security: {\it Security} and {\it
Malicious Code}. Our early findings indicated that from the total number of the
artifacts, 4,270 of them contained at least one bug coming from the first category
and 45,145 from the second.

Figure~\ref{fig:bug-per} shows how software bugs are distributed among the
repository. Together with the {\it Bad Practice} bugs and the {\it Style} bugs,
security bugs (the sum of the {\it Security} and {\it Malicious Code}
categories) are the most popular in the repository.

5,601 - jar size - 17,505. Total Projects: 5601, Versions: 32689.

>>>>

First, we scanned the Maven repository for appropriate {\sc jar}s and created a
list that included them. We discuss the {\sc jar} selection process in the next 
section. With the {\sc jar} list at hand, we created a series of processing tasks
and added them to the task queue. Then we executed twenty five (Unix-based)
workers that checked out tasks from the queue, processed the data and stored the
results to the data repository.

A typical processing cycle of a worker included the following steps: after
the worker spawns, it requests a task from the queue. This task contains
the {\sc jar} name, which is typically a project version that is downloaded locally.
First, specific {\sc jar} metadata are calculated and stored. Such metadata include
its size, its dependencies, and a number that represents the chronological order of the
release. This order is derived from an {\sc xml} file that
accompanies every project in the Maven repository called {\it
maven-metadata.xml}. Then FindBugs is executed and its results are also stored
in the data repository. When the task is completed the queue is notified and
the next task is requested.

@@

Initially, we obtained a snapshot (January 2012) of the maven repository and
handled it locally to retreive a list of all the names of the project versions
that existed in it. Since FindBugs analyzes applications written in the Java
programming language, and the Maven repository
hosts projects from languages other than Java such as Scala, Groovy,
Clojure etc. we filtered out such projects by performing a series of checks in
the repository data and metadata.

In addition, we implemented a series of audits in the worker scripts that
checked if the {\sc jar}s are valid in terms of implementation. For instance,
for every {\sc jar} the worker checked if there were any {\it .class} files
before invoking FindBugs. After the project filtering we narrowed down
our data set to 17,055 projects with 114,399 versions.
Table~\ref{tbl:repository} summarises the data set information and
provides the basic descriptive statistic measurements. The distribution of version
count among the selected projects is presented in Figure \ref{fig:version-count}.

@@

The statistical mesurements presented in Table~\ref{tbl:repository}
indicate that we have 17,505 projects total and the data sets median
is 3, which means that almost 50\% (8,753 projects) of the project
population has 1 to 3 versions.
Note that the Maven repository has many projects with a
few number of versions. For instance, there are numerous projects with a
number of versions that is less than ten.
There are a few projects containing ten
versions and only a few with hundreds of versions. The maximum number of
versions for a project is 337. The 3$^{rd}$ quartile measurement
also indicated that 75\% (13,129) of the projects have a maximum of 8 versions.

@@

\begin{table*}
\centering
\begin{center}
\leavevmode
	\begin{tabular}{l r}
	\hline
	Bug Description & Number of Artifacts\\
 	\hline
	{\sc hrs}: {\sc http} cookie formed from untrusted input & 149\\
	{\sc hrs}: {\sc http} response splitting vulnerability & 1,564\\
	{\sc pt}: absolute path traversal in servlet  & 92\\
	{\sc pt}: relative path traversal in servlet & 50\\
	{\sc sql}: nonconstant string passed to execute method on an {\sc sql} & 1,817\\
	{\sc sql}: a prepared statement is generated from a nonconstant String & 1,438\\
	{\sc xss}: {\sc jsp} reflected cross site scripting vulnerability & 17\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability in error page & 88\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability & 140\\
	\hline
	\end{tabular}
\caption{Number of Artifacts that Contain at Least One Severe Security Bug}
\label{tbl:sev}
\end{center}
\end{table*}

FindBugs separates software bugs into seven categories (see
Table~\ref{tbl:bug-cat}). Two of them involve security: {\it Security} and {\it
Malicious Code}. From the total number of the artifacts, 4,270 of them contained
at least one bug coming from the first category
and 45,145 coming from the second.

Another observation involves bugs that we could call {\it
severe}. Such bugs are related to vulnerabilities that appear due to the lack of user-input
validation and can lead to damaging attacks like {\sc sql} injection and
Cross-Site Scripting. Note that to exploit such vulnerabilities, a malicious user does
not have to know anything about the applications internals. For all the other
bugs, another program should be written either to incorporate references to
mutable objects, access non-final fields and others. In addition, as
bug descriptions indicate,\footnote{\url{http://findbugs.sourceforge.net/bugDescriptions.html}}
if an application has bugs like these, it might have more vulnerabilities that FindBugs doesn't
report. Table~\ref{tbl:sev} presents the number of artifacts where at least one of these
bugs exists. In the following section we provide results separately for this
subcategory.

Figure~\ref{fig:bug-per} shows how software bugs are distributed among the
repository. Together with the {\it Bad Practice} bugs and the {\it Style} bugs,
security bugs (the sum of the {\it Security} and {\it Malicious Code}
categories) are the most popular in the repository. In the following section we
show the correlation of security bugs with all the other categories. An interesting observation is
that 5,601 (31.99\%) projects, containing 32,689 artifacts (28.57\%), did not
have any bugs at all. These cases involved small artifacts with a minimum number of classes.
This consorts with the relationship between the size of an artifact and it's
bugs that is presented in the upcoming section.
